MSDS6372Project2

### 2021-04-06 MSDS6372Project2.Rmd
more update on mtry rf

#### Objective 2 part 3 - Nonparametric model approach - Random forest
- Default RF ntree = 500 and m variables =sqrt(p) seem to work better than ntree=100
- Try iterate through mtry variable from 1 to 10 using prediction cut off previously discovered - seems mtry = 1 works best on overall accuracy
- After actually fitting random forest with mtry = 1 the AUC = 0.882 (performance is worse compare to default mtry=3). Sensitivity and Specificity also suffers after adjusting the cut off. So Default RF option still wins!
- So Mtry = 3 seems to be the perfect balance for our dataset in order to provide good AUC.
```{r}
library(randomForest)
set.seed(1)
# By default use sqrt(p) number variables when building a random forest of classification trees
rf.adult = randomForest(income~.,data=adult)

@@

##### Find best mtry ####
# Use custom loop to find it based on training model
# Using previous discovered cutoff3 = 0.425 
parameter = 10
overall.Accuracy = c()
for(i in 1:parameter){
  set.seed(1)
  rf.adult.temp = randomForest(income~.,data=adult,mtry=i)
  rf.pred.temp = predict(rf.adult.temp,newdata=testdata,type = "prob")
  rf.prob.temp = rf.pred.temp[,2]
  class.rf.temp <-factor(ifelse(rf.prob.temp>cutoff3,">50K.","<=50K."),levels=c("<=50K.",">50K."))
  cm.temp = confusionMatrix(class.rf.temp,adult.test$income)
  overall.Accuracy[i]=cm.temp$overall[1]
}
overall.Accuracy

###### Giving us Mtry = 1 is the best split ####
# Try this option

>>>>

#### Objective 2 part 3 - Nonparametric model approach - Random forest
- Default RF ntree = 500 and m variables =sqrt(p) seem to work better than ntree=100
- Try iterate through mtry variable from 1 to 10 using prediction cut off previously discovered - seems mtry = 1 works best on overall accuracy
- After actually fitting random forest with mtry = 1 the AUC = 0.882 (performance is worse compare to default mtry=3). Sensitivity and Specificity also suffers after adjusting the cut off. So Default RF option still wins!
- So Mtry = 3 seems to be the perfect balance for our dataset in order to provide good AUC.
- We further added in calculations for AUC under each mtry #, and it also tells us mtry=3 provides the best AUC value
> overall.Accuracy
 [1] 0.8526934 0.8479922 0.8114267 0.7909892 0.7781913 0.7676787 0.7533790 0.7478289 0.7484166 0.7407770
> overall.auc
 [1] 0.882 0.903 0.907 0.903 0.899 0.893 0.890 0.887 0.883 0.878
```{r}
library(randomForest)
set.seed(1)
# By default use sqrt(p) number variables when building a random forest of classification trees
rf.adult = randomForest(income~.,data=adult)

@@

##### Find best mtry ####
# Use custom loop to find it based on training model
# Using previous discovered cutoff3 = 0.425 
# Adjust this to also find best mtry for AUC
parameter = 10
overall.Accuracy = c()
overall.auc = c()
for(i in 1:parameter){
  print(paste("Processing mtry=",i))
  set.seed(1)
  rf.adult.temp = randomForest(income~.,data=adult,mtry=i)
  rf.pred.temp = predict(rf.adult.temp,newdata=testdata,type = "prob")
  rf.prob.temp = rf.pred.temp[,2]
  class.rf.temp <-factor(ifelse(rf.prob.temp>cutoff3,">50K.","<=50K."),levels=c("<=50K.",">50K."))
  cm.temp = confusionMatrix(class.rf.temp,adult.test$income)
  overall.Accuracy[i]=cm.temp$overall[1]

  # Calculate AUC
  results.rf.temp <-prediction(rf.prob.temp, adult.test$income,label.ordering=c("<=50K.",">50K."))
  auc.temp <- performance(results.rf.temp, measure = "auc")
  auc.temp <- auc.temp@y.values
  overall.auc[i] <- round(auc.temp[[1]],3)
}
overall.Accuracy
overall.auc

###### Giving us Mtry = 1 is the best split based on overall.Accuracy ####
# Try this option
